---
title       : Assignment for Developing Data Products
subtitle    : Predicting Iris with Trees
author      : Joao Clemencio
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(GGally)
library(caret)
library(rattle)
```

## Iris analysis with a decision tree model

The Iris flower dataset is a famous multivariate data set that classifies three very similar species of plants using 4 measurements.

A classifier was built with a decision tree model to allow a user to interactively predict the correct species by selecting values for the 4 possible parameters:
* Sepal Length
* Sepal Width
* Petal Length
* Petal Width

--- .class #id 

## Exploratory analysis

Here is the relationships between the 4 considered variables, coloured by Species. Some distinct clusters can be observed.

```{r pairs, echo=FALSE, fig.align='center', fig.height=6, fig.width=13}
ggpairs(data = iris, columns = 1:4, mapping = aes(color = Species))
```

---

## Model creation

The model was created using the `rpart` method from the `caret` package.

```{r train, message = FALSE, warning=FALSE}
iris_rpart <- train(Species ~ ., data=iris, method="rpart")
print(iris_rpart$finalModel)
```

---

## Final decision tree

Here is the final decision tree

```{r tree, echo=FALSE, fig.align='center', fig.height=6, fig.width=13}
fancyRpartPlot(iris_rpart$finalModel)
```

--- 

## Prediction examples

```{r wat}
iris_mean = data.frame(lapply(iris[,-5], mean)); iris_sd = data.frame(lapply(iris[,-5], sd)); reps = 5
random_iris = data.frame(
  Sepal.Length = rnorm(reps, iris_mean$Sepal.Length, iris_sd$Sepal.Length),
  Sepal.Width  = rnorm(reps, iris_mean$Sepal.Width,  iris_sd$Sepal.Width),
  Petal.Length = rnorm(reps, iris_mean$Petal.Length, iris_sd$Petal.Length),
  Petal.Width  = rnorm(reps, iris_mean$Petal.Width,  iris_sd$Petal.Width))
random_iris$Species_prediction = predict(iris_rpart, random_iris)
print(random_iris)
```